---
title: "A Tale of Two Economies"
author: "Justin M. Shea"
date: "September 12, 2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(rmarkdown)
library(xts)
library(ggplot2)
library(blscrapeR)
library(tidyr)
library(ggthemes)
library(tufte)
library(RColorLisa)
library(dplyr)

source("functions.R")

library(RColorLisa)

Picasso <- color.lisa.pal(5, "The Dream by Pablo Picasso")
Basquiat <- color.lisa.pal(4, "DUSTHEADS by Jean-Michel Basquiat")
vangogh <- color.lisa.pal(4, "The Starry Night by Vincent van Gogh")
Basquiat <- color.lisa.pal(4, "La Mariée sous le Baldaquin by Marc Chagall")

```

## Introduction
```{r, message = FALSE, warning = FALSE, cache = TRUE}
df <- get_bls_county()
```

```{r, message = FALSE, warning = FALSE}
#Use map function with arguments.
map_bls(map_data = df, fill_rate = "unemployed_rate", 
        labtitle = "Unemployment Rate by County")
```

## Unemployment Rate

```{r message = FALSE, warning = FALSE, cache=TRUE}
# Employment Statistics
Unemployment_url <- "https://fred.stlouisfed.org/data/UNRATE.txt"

Unemployment <- as.xts(read.zoo(Unemployment_url, skip = 24, index.column = 1, 
                            header = TRUE, format = "%Y-%m-%d", FUN = as.yearmon))

Unemployment_chart <- data.frame(date = index(Unemployment["2000/"]), value = coredata(Unemployment["2000/"]))
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Series ID: UNRATE LNS14000000"}
ggplot(data = Unemployment_chart, aes(y = value, x = date)) +
        geom_line(color = Picasso[1], size = 1) +
        scale_y_continuous(position = "right") +
        xlab("Year") +
        ylab("Percent") +
        #geom_vline(xintercept = 2008.5, alpha = 1/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")
```




## Labor Force Participation Rate

```{r message = FALSE, warning = FALSE, cache=TRUE}
labor_participation_url <- "https://fred.stlouisfed.org/data/CIVPART.txt"

labor_participation <- as.xts(read.zoo(labor_participation_url, skip = 14, index.column = 1, 
                                   header = TRUE, format = "%Y-%m-%d", FUN = as.yearmon))

Participation_chart <- data.frame(date = index(labor_participation["1985/"]), value = coredata(labor_participation["1985/"]))
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Series ID: CIVPART LNS11300000"}
ggplot(data = Participation_chart, aes(y = value, x = date)) +
        geom_line(color = Picasso[1], size = 1) +
        scale_y_continuous(position = "right") +
        xlab("Year") +
         ylab("Percent") +
          geom_vline(xintercept = 2000, alpha = 1/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")
```

## The labor participation rate, by age

```{r message = FALSE, warning = FALSE, cache = TRUE}
#bls data by ages
library(blscrapeR)

employment_series <- c("LNS11324230", "LNS11300093", "LNS11300091", "LNS11300089", "LNS11324887" )
labels <- c("55 and over", "45 to 55", "35 to 44", "25 to 34", "16 to 24")

employment_ages <- bls_api(employment_series, startyear = 2000, end = 2017, registrationKey = Sys.getenv("BLS_KEY")) %>% 
        dateCast()

employment_wide <- employment_ages %>% spread(seriesID, value) %>% select(5:10) %>%
        arrange(date)
        
employment_diff <- apply(employment_wide[,2:6], 2, diff)

employment_sum <- apply(employment_diff, 2, cumsum)

employment_final <- data.frame(employment_wide[-1,"date"], employment_sum) %>%
         rename(".55 and over"=LNS11324230, "45 to 55" = LNS11300093, "35 to 44" = LNS11300091,
                "25 to 34" = LNS11300089, "16 to 24" = LNS11324887)
        
employment_long <- gather(employment_final, "Age", "Percent", 2:6)
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "LNS11324230, LNS11300093, LNS11300091, LNS11300089, LNS11324887"}
  ggplot(employment_long, aes(x=date, y=Percent)) +
        geom_line(aes(colour = Age), size = 1) +
        scale_y_continuous(position = "right") +
        xlab("Year") +
         ylab("Percent") +
        geom_hline(yintercept = 0, alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")

```

## Real Median Household Income

```{r message = FALSE, warning = FALSE, cache=TRUE}
# Employment Statistics
Income_url <- "https://fred.stlouisfed.org/data/MEHOINUSA672N.txt"

Income <- as.xts(read.zoo(Income_url, skip = 48, index.column = 1, 
                            header = TRUE, format = "%Y-%m-%d", FUN = as.yearmon))

Income_chart <- data.frame(date = index(Income["1985/"]), value = coredata(Income ["1985/"]))
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Series ID: MEHOINUSA672N"}
ggplot(data = Income_chart, aes(y = value, x = date)) +
        geom_line(color = Picasso[1], size = 1) +
        geom_smooth() +
        scale_y_continuous(position = "right") +
        xlab("Year") +
        ylab("2015 CPI-U-RS Adjusted Dollars") +
        geom_vline(xintercept = 2000, alpha = 1/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")
```

## Real GDP per capita VS Potential

```{r message = FALSE, warning = FALSE}
Real_GDP_URL <- "https://fred.stlouisfed.org/data/A939RX0Q048SBEA.txt"
Real_GDP <- as.xts(read.zoo(Real_GDP_URL, skip = 14, index.column = 1, 
                        header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))

HL_GDP_actual <- HLfilter(Real_GDP, h = 64)

Gap_Percent <- 100 * (sum(HL_GDP_actual["2017"]$HL_resid) / sum(HL_GDP_actual["2017"]$x_h))
#kable(Gap_Percent)

# log real gdp for plots
Hlm_GDP <- HLfilter.lm(Real_GDP["/2007"], h = 40)
HL_GDP <- HLfilter(Real_GDP, h = 40)

Potential <- HL_GDP$HL_fit["2008-10/"]

HL_output_xts <- merge(HL_GDP$x_h, Potential)["1985/"]

HL_output <- data.frame(index(HL_output_xts), coredata(HL_output_xts))
colnames(HL_output) <- c("date", "Actual", "Potential")

HL_long <- gather(HL_output, "Series", "Dollars", 2:3)
# plot GDP
#plot(HL_GDP$HL_fit, col = "red", grid.col = "white")
#lines(HL_GDP$x_h, col = "black")
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Series ID: A939RX0Q048SBEA, BEA Account Code: A939RX0"}
  ggplot(HL_long, aes(x = date, y = Dollars)) +
        geom_line(aes(colour = Series), size = 1) +
        scale_y_continuous(position = "right") +
         xlab("Year") +
         ylab("2009 Dollars") +
        geom_vline(xintercept = 2008.75, alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")



#plot(HL_GDP$HL_fit["1984/"], col = "red", grid.col = "white")
#lines(HL_GDP$x_h["1984/"], col = "black")
```

```{r, message=FALSE, warning=FALSE, eval=FALSE}
# plot resid
plot(HL_GDP$HL_resid["1952/"], index(employ_hlf), col = "black", grid.col = "white")
abline(v = .index(HL_GDP["2001"])[1], lty = 2, col = "green")
abline(v = .index(HL_GDP["2009"])[1], lty = 2, col = "red")
#abline(h = 415, lty = 2, col = "green")
abline(h = 0, lty = 2, col = "black")
#abline(h= -515, lty = 2, col = "red")
```


## Declining Government Spending
```{r, message = FALSE, warning = FALSE, tidy = TRUE}

### Test Fisher Tranformation Method using GDP 
# Define file paths
GDP_nom <- "https://fred.stlouisfed.org/data/GDP.txt"
GDP_Quantity_Index <- "https://fred.stlouisfed.org/data/A191RA3Q086SBEA.txt"
GDP_Price_Index <- "https://fred.stlouisfed.org/data/A191RG3Q086SBEA.txt"


### Government Consumption Expenditures
                        Federal_GCE <- "https://fred.stlouisfed.org/data/FGCE.txt"
                    State_Local_GCE <- "https://fred.stlouisfed.org/data/SLCE.txt"
                    
## GCE Quantity & Price Indexes
Federal_Quantity_Index <- "https://fred.stlouisfed.org/data/B823RA3Q086SBEA.txt"
        Federal_Price_Index <- "https://fred.stlouisfed.org/data/B823RG3Q086SBEA.txt"
#-------------------------------------#
State_Local_Quantity_Index <- "https://fred.stlouisfed.org/data/B829RA3Q086SBEA.txt"
        State_Local_Price_Index <- "https://fred.stlouisfed.org/data/B829RG3Q086SBEA.txt"
        
GCE_Table <- t(data.frame(Federal_GCE, Federal_Quantity_Index, Federal_Price_Index, State_Local_GCE, State_Local_Quantity_Index, State_Local_Price_Index))
#kable(GCE_Table, align = "l", caption = "GCE Nominal, Quantity, and Price index sources.")
```

```{r, message = FALSE, warning = FALSE, tidy = TRUE}
# Get data
GDP          <- as.xts(read.zoo(GDP_nom, skip = 19, index.column = 1, header = TRUE, 
                       format = "%Y-%m-%d", FUN = as.yearqtr))
GDP_Quantity <- as.xts(read.zoo(GDP_Quantity_Index, skip = 14, index.column = 1, 
                        header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
GDP_Price    <- as.xts(read.zoo(GDP_Price_Index, skip = 14, index.column = 1, 
                             header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
#------------------------------# 

Federal          <- as.xts(read.zoo(Federal_GCE, sep = "", skip = 13, 
                           index.column = 1,  header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
Federal_Quantity <- as.xts(read.zoo(Federal_Quantity_Index, sep = "", skip = 13, 
                           index.column = 1,  header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
Federal_Price    <- as.xts(read.zoo(Federal_Price_Index  , sep = "", skip = 13, 
                           index.column = 1,  header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
#------------------------------# 
State_Local          <- as.xts(read.zoo(State_Local_GCE, sep = "", skip = 13, 
                              index.column = 1,  header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
State_Local_Quantity <- as.xts(read.zoo(State_Local_Quantity_Index, sep = "", skip = 13, 
                              index.column = 1,  header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
State_Local_Price    <- as.xts(read.zoo(State_Local_Price_Index  , sep = "", skip = 13, 
                              index.column = 1,  header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
```


```{r message = FALSE, warning = FALSE, tidy = TRUE}
rGDP <- Fisher(GDP, GDP_Price, GDP_Quantity, 249) 
Real_Federal      <- Fisher(Federal, Federal_Price, Federal_Quantity, 249)
Real_State_Local  <- Fisher(State_Local, State_Local_Price, State_Local_Quantity, 249)
```

```{r message = FALSE, warning = FALSE}

Government <- merge(Real_Federal,Real_State_Local)["1985/"]


Government_df <- data.frame(index(Government), coredata(Government))
colnames(Government_df) <- c("date", "Federal", ".State & Local")

Government_long <- gather(Government_df, "Series", "Spending", 2:3)
# plot GDP
#plot(HL_GDP$HL_fit, col = "red", grid.col = "white")
#lines(HL_GDP$x_h, col = "black")
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Real Federal and Local Spending, 1985-Present"}
  ggplot(Government_long , aes(x = date, y = Spending)) +
        geom_line(aes(colour = Series), size = 1) +
        scale_y_continuous(position = "right") +
         xlab("Year") +
         ylab("Billions of 2009 Dollars") +
        geom_vline(xintercept = 2008.75, alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")

#plot(HL_GDP$HL_fit["1984/"], col = "red", grid.col = "white")
#line
```

```{r}
# GDP
GDP_and_Gov <- merge(Real_GDP, Real_Federal, Federal, Real_State_Local, State_Local)
Government_table <- data.frame(coredata(GDP_and_Gov ["2010/"]))
rownames(Government_table) <- index(GDP_and_Gov["2010/"]) 
#kable(Government_table)


total_table <- data.frame(sapply(Government_table, function(x) (x[NROW(x)] - x[1])))
summary_table <- t(total_table)
rownames(summary_table) <- "Change from 2010"
#kable(summary_table)
```

```{r , message = FALSE, warning = FALSE}
Federal_GDP_percent <- 100 * (Real_Federal/rGDP)
State_GDP_percent <- 100 * (Real_State_Local/rGDP)
Government_GDP <- merge(Federal_GDP_percent, State_GDP_percent)["1985/"]

Government_GDP_df <- data.frame(index(Government_GDP), coredata(Government_GDP))
colnames(Government_GDP_df) <- c("date", "Federal", ".State & Local")

Government_GDP_long <- gather(Government_GDP_df, "Series", "Spending", 2:3)
# plot GDP
#plot(HL_GDP$HL_fit, col = "red", grid.col = "white")
#lines(HL_GDP$x_h, col = "black")
```

```{r, eval=FALSE, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Government Expenditure: Percentage of Real GDP"}
  ggplot(Government_GDP_long, aes(x = date, y = Spending)) +
        geom_line(aes(colour = Series), size = 1) +
        scale_y_continuous(position = "right") +
         xlab("Year") +
         ylab("Percent of GDP") +
        geom_vline(xintercept = c(2008.75), alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")

#plot(HL_GDP$HL_fit["1984/"], col = "red", grid.col = "white")
#line
```


## Long Term Corporate Earnings


```{r message = FALSE, warning = FALSE, cache=TRUE}

library(Quandl)
Quandl.api_key("wzAyVxVa7H7Rj1xZqy6V")
SP500_E <-  Quandl("MULTPL/SHILLER_PE_RATIO_MONTH", type="xts")
SP500_Sales <- Quandl("MULTPL/SP500_REAL_SALES_YEAR", type="xts")
SP500_Earnings <- Quandl("MULTPL/SP500_EARNINGS_YEAR", type="xts")

SP500_Profit <- (SP500_Sales-SP500_Earnings)/SP500_Earnings

SP500_E_chart <- data.frame(date = index(SP500_E["1920/"]), value = coredata(SP500_E["1920/"]))
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Robert Shiller: Cyclically Adjusted PE Ratio (CAPE Ratio)"}
ggplot(data = SP500_E_chart, aes(y = value, x = date)) +
        geom_line(color = Picasso[1], size = 1) +
        scale_y_continuous(position = "right") +
        xlab("Year") +
        ylab("Earnings Per Share") +
        geom_vline(xintercept = c(1929, 2000, 2017), color = "red", alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")

```

## S&P 500 Revenue Per Share


```{r message = FALSE, warning = FALSE, cache=TRUE}

library(Quandl)
Quandl.api_key("wzAyVxVa7H7Rj1xZqy6V")

SP500_Sales <- Quandl("MULTPL/SP500_REAL_SALES_QUARTER", type="xts")

SP500_E_chart <- data.frame(date = index(SP500_Sales), value = coredata(SP500_Sales))
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Real 12-Month S&P 500 Sales Per Share"}
ggplot(data = SP500_E_chart, aes(y = value, x = date)) +
        geom_line(color = Picasso[1], size = 1) +
        scale_y_continuous(position = "right") +
        xlab("Year") +
        ylab("Earnings Per Share") +
        geom_vline(xintercept = c(2008), color = "red", alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")

```


```{r, message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Series ID: CP,  BEA Code: A055RC1"}

GDP_nom <- "https://fred.stlouisfed.org/data/GDP.txt"
GDP          <- as.xts(read.zoo(GDP_nom, skip = 19, index.column = 1, header = TRUE, 
                       format = "%Y-%m-%d", FUN = as.yearqtr))


### Income Savings ###
Corporate_Profits_After_Tax <- "https://fred.stlouisfed.org/data/CP.txt"
Corporate_Net_Cash_Flow     <- "https://fred.stlouisfed.org/data/CNCF.txt"

### Download 

Corporate_Profits   <-  as.xts(read.zoo(Corporate_Profits_After_Tax, sep = "", skip = 10, 
                           index.column = 1,  header = TRUE, format = "%Y-%m-%d", 
                           FUN = as.yearqtr))
Corporate_Cash_Flow <-  as.xts(read.zoo(Corporate_Net_Cash_Flow, sep = "", skip = 13, 
                           index.column = 1,  header = TRUE, format = "%Y-%m-%d", 
                           FUN = as.yearqtr))     

Corporate_GDP_percent <- 100 * (Corporate_Profits/GDP)

#autoplot(Corporate_GDP_percent["1960/"]) +
 #       ylab("Corporate_Profits, % GDP") +
  #      xlab("Year") +
   #     scale_y_continuous(position = "right") +
    #    geom_vline(xintercept = c(2008)) +
     #   theme_tufte()
```

## Corporate Profits, Percentage of Cash Flow

```{r, message = FALSE, warning = FALSE}

Corporate_Net_Cash_Flow     <- "https://fred.stlouisfed.org/data/CNCF.txt"


Corporate_Cash_Flow <-  as.xts(read.zoo(Corporate_Net_Cash_Flow, sep = "", skip = 13, 
                           index.column = 1,  header = TRUE, format = "%Y-%m-%d", 
                           FUN = as.yearqtr))     

Corp_Cash_GDP_percent <- 100 * (Corporate_Profits/Corporate_Cash_Flow)

Corp_Cash_GDP_chart <- data.frame(date = index(Corp_Cash_GDP_percent["1982/"]), value = coredata(Corp_Cash_GDP_percent["1982/"]))
```

```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Series: CP & CNCF"}
ggplot(data = Corp_Cash_GDP_chart, aes(y = value, x = date)) +
        geom_line(color = Picasso[1], size = 1) +
        scale_y_continuous(position = "right") +
        xlab("Year") +
        ylab("Corporate Profits, % of Corporate Cash Flow") +
        geom_vline(xintercept = c(2008), color = "red", alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")

```



```{r, eval=FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 3, fig.fullwidth = TRUE, fig.cap = "Corporate Cash Flows without profits: Percentage of Real GDP"}
cash_profit <- 100 * (Corporate_Cash_Flow-Corporate_Profits)/GDP

autoplot(cash_profit["1982/"]) +
        ylab("Corporate Profits, % of Corporate Cash Flow") +
        xlab("Year") +
        scale_y_continuous(position = "right") +
        geom_vline(xintercept = c(2008)) +
        theme_bw()

```




## Real Minimum Wage, in 2009 Dollars

```{r, message = FALSE, warning = FALSE, cache = TRUE}

minimum_oath <- "https://fred.stlouisfed.org/data/FEDMINNFRWG.txt"
minimum_wage <-  as.xts(read.zoo(minimum_oath, sep = "", skip = 26, 
                           index.column = 1,  header = TRUE, format = "%Y-%m-%d", FUN = as.yearmon)) 

minimum_wage_qtr <- to.quarterly(minimum_wage)["1947/"]

minimum_wage_yr <- to.yearly(minimum_wage)["1947/"]

GDP_Quantity_Index <- "https://fred.stlouisfed.org/data/A191RA3Q086SBEA.txt"
GDP_Price_Index <- "https://fred.stlouisfed.org/data/A191RG3Q086SBEA.txt"
# Get data
GDP_Quantity <- as.xts(read.zoo(GDP_Quantity_Index, skip = 14, index.column = 1, 
                        header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))
GDP_Price    <- as.xts(read.zoo(GDP_Price_Index, skip = 14, index.column = 1, 
                             header = TRUE, format = "%Y-%m-%d", FUN = as.yearqtr))

base_year <- inflation_adjust(base_year=2009)

Real_minimum_wage <- minimum_wage_yr [,4]/base_year$adj_value
```

```{r}
colnames(Real_minimum_wage) <- "value"

minimum_wage_chart <- data.frame(date = index(Real_minimum_wage), value = coredata(Real_minimum_wage))
#Real_minimum_wage <- Fisher(minimum_wage_qtr[,4], GDP_Price, GDP_Quantity, base = 282)
#Real_minimum_wage
```


```{r message = FALSE, warning = FALSE, fig.width = 7, fig.height = 5, fig.fullwidth = TRUE, fig.cap = "Series: FEDMINNFRWG"}
ggplot(data = minimum_wage_chart, aes(y = value, x = date)) +
        geom_line(color = Picasso[1], size = 1) +
        scale_y_continuous(position = "right") +
        xlab("Year") +
        ylab("Dollars Per Hour") +
        geom_vline(xintercept = c(1969, 2009), color = "red", alpha = 2/3, linetype = "dashed") +
        scale_color_manual(values = Picasso) +
        theme_tufte(base_size = 16) + theme(legend.position = "top")

```





## Card and Krueger: Minimum Wages and Employment

```{r, message = FALSE, warning = FALSE, tidy = TRUE}
## Card and Krueger
library(wooldridge)
data("discrim")

discrim$NewJersey <- as.factor(discrim$NJ)
#str(discrim)
#lm(discrim$emp ~ discrim$wagest)
#plot(x=discrim$emp, y = discrim$wagest)

#ggplot(data=discrim, aes(x=emp, y=wagest)) + 
  #      geom_point() +
   #     geom_smooth(method = "lm")

ggplot(data=discrim, aes(x=emp, y=wagest, col = NewJersey)) + 
        geom_count(aes(size = ..prop.., group = NewJersey), alpha = 1/3) +
        scale_size_area(max_size = 10) +
        xlab("Number of Employees") +
        ylab("Starting Wage, Dollars") +
        ggtitle("February 1992 - Pre Minimum Wage Increase") + 
        scale_color_manual(values = Picasso) +
        theme_bw(base_size = 16) 

```

## Card and Krueger: Minimum Wages and Employment 2


```{r, message = FALSE, warning = FALSE, tidy = TRUE}

ggplot(data=discrim, aes(x=emp2, y=wagest2, col = NewJersey)) + 
         geom_count(aes(size = ..prop.., group = NewJersey), alpha = 1/3) +
        scale_size_area(max_size = 10) +
        xlab("Number of Employees") +
        ylab("Starting Wage, Dollars") +
        ggtitle("November 1992 - Post Minimum Wage Increase") + 
        scale_color_manual(values = Picasso) +
        theme_bw(base_size = 16)


```



